Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem \thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}

### 2018-05-21 Response.tex

\begin{document}

%\pagestyle{empty}

\section*{Second round of responses to the reviewers' comments for JASS-D-17-00005}

We thank the reviewers for their comments and aid in improving the quality of our manuscript. 
In accordance with the comments and suggestions, the manuscript has been revised, and the answers to all comments are addressed as follow.

@@

\subsection*{Reviewer 1}
\begin{itshape}
This is a vastly improved iteration of the paper and I thank the authors for taking the time to improve it.  I commend the authors for responding to all of the feedback, especially modifying figures such as Figure 2, which I know takes a lot of time.  I was also happy to see the intentional reductions in the paper to make it more focused, and the fact that the authors reconciled their low-thrust transfers with existing technology.  I still have a few issues that persist based on the current submission, but the paper is far closer to being acceptable for publication.
\end{itshape}

\begin{itemize}
    \item 
        \begin{itshape}
            Section 2.3

            "This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge-Kutta methods."

@@

            "Potentially lower state accuracy" is still overselling.  Just because your second order method is bounded on an energy surface does not guarantee that the state is accurate.  It is likely that a 2nd order integrator in the PCRTBP contains more state error than a high-order RK which is the standard for modern trajectory planning in flight operations.  I think it's best to focus on the primary advantage that the variational integrator provides reduced computational demands.
        \end{itshape}

    The manuscript has been modified to remove the about the benefits of state accuracy.
    The section has been modified to highlight the reduced computational load of the variational integrator.
    In addition, an additional sentence has been added specifically highlighting the fact that high order Runge-Kutta methods will probably provide better state accuracy at the expense of additional computations and poorer energy behavior.

    \begin{correction}
        This work uses a second order integrator which has improved energy behavior in comparison to first order integration techniques without a significant increase in computational demand.
        The variational integrator is constructed according to discrete Hamilton's principle, specifically for the given dynamic system, and it provides a long-term structural stability in capturing the effects of low-thrust propulsion systems. 
        General purpose numerical integrators, such as adaptive stepsize Runge-Kutta methods may yield a more accurate state trajectory depending on the choice of integration parameters, but at the expense of additional computational demand and total energy deviations.

@@

        \begin{correction}
            In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm with a convergence tolerance of \num{1e-5}.
            Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
            All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
            The numerical results are specific to the choice of nonlinear solver, and different tolerances or software tools may result in slight changes.
        \end{correction}
    \item \begin{itshape}
Conclusions:

"The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints."

I don't agree with the statement. Often, indirect methods are exceptionally difficult to append path constraints onto.  In contrast, direct methods such as collocation approaches, are very amenable to the formulation of path constraints.
    \end{itshape}

    Thank you for the comments. This sentence has been modified.
    \begin{correction}
        The indirect optimal control formulation directly solves for the state and costate vectors in the form of a two point boundary value problem which results in a lower number of design variables.
        Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution and typically results in a much larger parameter set.
    \end{correction}
\end{itemize}
\subsection*{Reviewer 2}

@@

\begin{itshape}
ISSUE 1: introduction (requires minor revisions): 
Despite some reduction, the intro is still long-winded.  For example,
\end{itshape}

The introduction has been reduced in accordance with the recommendations below.
Several paragraphs have been removed or refactored to reduce repetition and extraneous detail.
The updated introduction is included below and specific comments for each comment is also provided.

\begin{itemize}

    \item
        \begin{itshape}
            It's superfluous to explain at such length that spacecraft trajectory design is a problem of interest, that CubeSats are a thing, etc. Broad points like these could potentially be condensed into single, citation-heavy sentences. 
        \end{itshape}

@@

            %%-------------- Motivation for low thrust orbital transfers

            % enables long duration missions with minimized propellant usage since Isp is so high

            % Enables different mission scenarios which are not possible through the use of conventional chemical propulsion (high thrust) systems
            Designing spacecraft trajectories is a classic and ongoing topic of research.
            There has been significant research into the design of orbital transfers for space vehicles.
            Optimal expenditure of onboard propellant is critical to allowing a mission to continue for a longer period of time or to enable the launch of a less massive spacecraft.
            Electric propulsion systems offer a much greater specific impulse than chemical systems.
            As a result, the greatly increased efficiency allows for greater payload mass or extended duration missions.
            However, these electric propulsion systems typically have much less thrust than their chemical counterparts and therefore orbital maneuvers have a much longer time of flight.
            In spite of this drawback, a wide variety of missions, such as communication and deep space probes, have utilized the unique benefits of low thrust electric propulsion to great effect~\cite{choueiri2009}.

            % utilizing the combo of small satellite and low thrust propulsion offers new mission scenarios 
            Recent developments in miniature electric propulsion and small satellites now offer the potential for new research opportunities~\cite{folta2015,haque2013}.
            The potential for more demanding missions places an even greater importance on the mission design to ensure that optimal trajectories satisfy mission requirements~\cite{folta2015,koon2011,ross2006,gomez2001}. 
            There has been extensive research focused on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011,koon2011,ross2006}. 
            Within the three-body problem, a spacecrafts feasible region of motion is constrained by its energy, or Jacobi integral.
            The addition of low-thrust propulsion offers the potential of reduced transfer transit times and the ability to depart from the free motion trajectory to allow for increased transfer opportunities.
            Frequently, insight into the problem or intuition on the part of the designer is required to determine initial conditions that will converge to the desired solution due to the inherent nonlinear and chaotic behavior of the three-body system~\cite{szebehely1967}.
            Conventional general-purpose Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, may fail to preserve geometric properties of the dynamic system numerically. 
            Variational integrators can allow for reduced computational effort without any loss of numerical stability or energy drift which exist in conventional integration schemes.

@@

            In short, the authors present a systematic method of generating optimal transfer orbits in the three-body problem.
            This paper provides a discrete optimal control formulation to generate the reachability set on a \Poincare section.
            % This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods which rely on coarse or suboptimal control parameterizations.
            % For example, a hybrid direct method may use a continuous control law developed from indirect optimal control theory while directly minimizing a cost function instead of solving the full two-point boundary value problem~\cite{ozimek2010a}.
            In addition, the use of a geometric integrators ensures numerical stability for long-duration orbit transfers and maintains this behavior with the addition of small magnitude control inputs.
            Our computation of the reachable set allows for a simple metric of defining optimal trajectories.
            We avoid the issues inherent in selecting a valid initial condition for optimization.
            Rather, we choose a state on the reachability set which minimizes the distance toward the desired target.
            We demonstrate these capabilities via two numerical examples simulating transfer trajectories in the Earth-Moon system.

        \end{correction}
\end{itemize}

\begin{itshape}
ISSUE 2: analysis/illustration (requires major revisions): 
My other highlighted area of concern with the original manuscript was the degree of (in)completeness in illustrating the method and its pros and cons.  I do not feel that the authors have done very much to improve this aspect of the paper beyond adding some caveats that feel a little half-hearted rather than properly contextualized discussions.  This aspect of the research remains a major shortcoming that must be addressed before publication.

While my request for improvements was itself not very specific, I had hoped that some of my other questions would reveal some areas of consideration for the authors.  Addressal/discussion of those questions furthermore needs to appear adequately in the paper, not just the rebuttal. Below I will try to sketch how these points of inquiry can lead to a more demonstrative analysis.
\end{itshape}
\newline\newline
Thank you for the comments and the additional suggestions to improve the manuscript. 
The main objective of the submitted manuscript is to construct a systematic approach to design low-thrust orbital maneuvers. 
It is motivated by the specific challenges that electric propulsion systems generate minuscule level of thrust, and low-thrust orbital maneuvers rely on those effects accumulated over a long time period. 
It is quite cumbersome and challenging to apply numerical optimization tools or boundary value problem solvers successfully to low-thrust orbital design problems. 
We aim to provide more organized tools to mitigate such difficulties by explicitly considering the characteristics of the natural dynamics and utilizing geometric integrators exhibiting long-term numerical stability. 
The former is achieved by incorporating the control-free maneuvers obtained by the Poincare section in the invariant manifolds, and the latter is achieved by variational integrators. 
However, the proposed scheme does not guarantee any optimality in the constructed maneuvers. 
It is heuristically implied that the constructed orbital maneuver would require no-excessive control input, as it is essentially constructed by adjusting the control-free trajectory. 
In short,
\begin{itemize}
	\item[Pros] The proposed scheme utilizes the characteristics of the inherent three-body problem, in constructing low-thrust orbital maneuvers in a systematic way. 
		This reduces the cumbersome process of guessing a \textit{good} set of initial parameters to make the direct application of optimization become successful. 
	\item[Pros] The use of geometric integrators is particularly useful in capturing the long-term effects of low thrust computationally effectively. 
	\item[Cons] While the proposed scheme generates orbital trajectories to the target with prescribed constraints, it does not guarantee any optimality. 
\end{itemize}
The conclusions section has been revised to summarize the above pros and cons of the proposed approach. 
\begin{itshape}
	Put the revised conclusions here.
\end{itshape}

Also, the revised manuscript incorporates many of the suggested changes and additions.
We include several new figures to further illustrate the change in Jacobi energy.
In addition, we add some additional examples demonstrating the effect of variations in the control bound and terminal time on the reachability set.

@@

\begin{itemize}
    \item 
        \begin{itshape}
            The assumption of fixed travel time.  The authors respond that this choice is made to simplify the optimization problem.  Of course this is the case, and is conventional, but this choice is made in a rather arbitrary fashion rather than properly analyzed (especially considering that a central purpose of the problem setup is to enable faster travel times).  Imagine plotting low-thrust reachability curves as in Fig8b for each of a sequence of travel time choices.  This would reveal the influence of the arbitrarily handled parameter, and show that there's some critical point at which the reachablility curve intersects the targeted manifold.  Maybe exhaustive exploration of this dimension is not appropriate in every application, but the relationship is worth illustrating.
        \end{itshape}

        Thank you for the comment. 
        We agree that some further discussion of the effect of these ``tunable'' parameters would help to improve the manuscript. 
		According to the comments, we considered twelve additional cases for every combination of varying terminal time $t_f$ and the maximum control $u_m$ listed at Table \ref{fig:varying_tf_reachability_sets}. 
        An additional section has been included which highlights some of the effects of variations in terminal time, and control magnitude (next comment).
        The results presented are not an exhaustive description of the search space (in terms of travel time and/or control magnitude) but do highlight some of the important characteristics of these parameters on the resulting reachability set.
        The additional discussion and plots are shown below.

@@

    \subcaptionbox{Jacobi Energy: Energy history throughout the transfer.
    Each stage of the transfer serves to increase the energy towards the stable manifold. 
The final reachability set intersects the stable manifold and enables a large final manuever to the target orbit.\label{fig:jacobi_geo}}{\includegraphics[width=0.4\textwidth]{figures/geo_transfer/jacobi_energy.pdf}}

        \caption{Geostationary to \( L_1 \) periodic orbit transfer: Complete transfer from the geostationary orbit to the stable manifold.\label{fig:geo_transfer}}
\end{figure}
        \end{correction}
    \item 
        \begin{itshape}
            Following that thought, I have serious concerns about chaining this Poincare-based approach, as done in the second problem and as I previously asked about.  The authors' responded: "While the trajectories do return to the energy level of the Poincare section, there is a transfer of "energy" between the states of the vehicle." Frankly, this statement doesn't make sense to me.  I think that you are actually getting a very inefficient solution by forcing these repeated (and at fixed intervals) returns to the same energy level (again, plot the Jacobi time series and see if you can square it with your intuition).  Furthermore posing the problem as a transfer to the stable manifold kind of hamstrings the motivation of reducing transfer times.  Honestly, I think this second example should simply be axed.  If the first one is expanded enough to illustrate the problem fundamentals, it will be sufficient.  Conducting the same kind of analysis at a couple different rev counts would
            reveal the dimension that this second example added, and help reveal the limits of the approach under certain scenarios and thrust capabilities.
        \end{itshape}

		As pointed out by the reviewer, our respond stating `..the trajectories do return to the energy level of the Poincare section..' is not correct. 
        Instead, the combination of several reachability sets does not force the system to return to a prescribed energy level.
        Each reachability set computation serves to maximize a distance metric, defined in a reduced phase space corresponding to the \Poincare section.
        The cost function is computed as the distance between the terminal states of the controlled and uncontrolled trajectories.
        The additional plots shown in both examples highlight the fact that the Jacobi energy is continuously changing as the control input is used to maneuver the vehicle towards the desired target states.

@@

        The transfer to the stable manifold is presented as an example of the combination of existing control free  transfer methodologies with low thrust control inputs.
        This example demonstrates one possible use of low-thrust control to enable a larger region of feasibility for the use of invariant manifold transfers in the three body problem.

		With regards to the comments on the second example, the authors decided to keep them in the manuscript, as it illustrates the idea of multiple stages to construct orbital maneuvers with more challenging scenarios. 
		Nonetheless, the revised manuscript includes extended numerical results and discussions for the first example, for example to illustrate the effects of varying $t_f$ and $u_m$ and the evolution of the Jacobi constant. 
    \item 
        \begin{itshape}
            To summarize my thoughts on ISSUE 2:
            As they exist, the examples illustrate that solutions can be found but they do not enlighten the reader as to the characteristics of the problem formulation/tool that should be understood and kept in mind when applying it or when planning related research efforts.  Furthermore, I am very suspicious that Example 2 is not actually a good solution or a compelling use case for the approach.  I do not necessarily ask or expect the authors to follow my above outline precisely, but I insist that some changes occur to make the analysis/illustrations better convey the general nature of the tool, not just that it gives you an answer.  
        \end{itshape}

@@

        Thank you for the comments.
        The manuscript has been editied, as described previously, to incorporate several changes to illustrate the specific characteristics of this approach.
        An additional section is presented which highlights the effect of terminal time and control magnitude on the reachability set.

        The second example presents a larger and more involved example of combining several iterations of the reachable set to achieve a larger and more complex transfer. 
        This aspect of the proposed method, more specifically of combining several reachability sets to approach a target, is not demonstrated in the first example.
        As a result, we feel that the second example does add some additional insight into the use of this method for a more complicated transfer.
		Furthermore, the revised conclusions section summarizes the objective of the submitted manuscript, and the pros and the cons of the proposed approach.

@@

    \item
        \begin{itshape}
            A more collected and focused discussion about how the parameter choices, assumptions, and limitations of the tool should inform its use is also warranted; currently remarks to this effect feel like scattered afterthoughts.  One organized way to think about this is to distinguish all the dimensions of the problem and identify what is being solved rigorously, what is being set manually, what is being numerically explored, etc.
        \end{itshape}

        Thank you for the helpful comments to improve the manuscript.
        The manuscript has been edited to reduce the length and repetition of the introduction.
        In addition, an additional section is presented highlighting the effect of the variations of terminal time and control magnitude on the reachable set. 
        Furthermore, the presented examples also include additional plots showing the change of the Jacobi energy during the transfer. 
\end{itemize}
% \bibliography{library}
% \bibliographystyle{spmpsci}
\end{document}

### 2018-05-21 Response.tex

            "This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge-Kutta methods."

            "Potentially lower state accuracy" is still overselling.  Just because your second order method is bounded on an energy surface does not guarantee that the state is accurate.  It is likely that a 2nd order integrator in the PCRTBP contains more state error than a high-order RK which is the standard for modern trajectory planning in flight operations.  I think it's best to focus on the primary advantage that the variational integrator provides reduced computational demands.
        \end{itshape}

    The manuscript has been modified to remove the about the benefits of state accuracy.
    The section has been modified to highlight the reduced computational load of the variational integrator.
    In addition, an additional sentence has been added specifically highlighting the fact that high order Runge-Kutta methods will probably provide better state accuracy at the expense of additional computations and poorer energy behavior.

    \begin{correction}
        This work uses a second order integrator which has improved energy behavior in comparison to first order integration techniques without a significant increase in computational demand.
        The variational integrator is constructed according to a discrete version of Hamilton's principle, specifically for the given dynamic system, and it provides long-term structural stability in capturing the effects of low-thrust propulsion systems.
        General purpose numerical integrators, such as adaptive stepsize Runge-Kutta methods may yield a more accurate state trajectory depending on the choice of integration parameters, but at the expense of additional computational demand and total energy deviations.

@@

            In short, the authors present a systematic method of generating optimal transfer orbits in the three-body problem.
            This paper provides a discrete optimal control formulation to generate the reachability set on a \Poincare section.
            % This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods which rely on coarse or suboptimal control parameterizations.
            % For example, a hybrid direct method may use a continuous control law developed from indirect optimal control theory while directly minimizing a cost function instead of solving the full two-point boundary value problem~\cite{ozimek2010a}.
            In addition, the use of a geometric integrators ensures numerical stability for long-duration orbit transfers and maintains this behavior with the addition of small magnitude control inputs.
            Our computation of the reachable set allows for a simple metric of defining optimal trajectories.
            We avoid the issues inherent in selecting a valid initial condition for optimization.
            Rather, we choose a state on the reachability set which minimizes the distance toward the desired target.
            We demonstrate these capabilities via two numerical examples simulating transfer trajectories in the Earth-Moon system.

        \end{correction}
\end{itemize}

\begin{itshape}
ISSUE 2: analysis/illustration (requires major revisions): 
My other highlighted area of concern with the original manuscript was the degree of (in)completeness in illustrating the method and its pros and cons.  I do not feel that the authors have done very much to improve this aspect of the paper beyond adding some caveats that feel a little half-hearted rather than properly contextualized discussions.  This aspect of the research remains a major shortcoming that must be addressed before publication.

While my request for improvements was itself not very specific, I had hoped that some of my other questions would reveal some areas of consideration for the authors.  Addressal/discussion of those questions furthermore needs to appear adequately in the paper, not just the rebuttal. Below I will try to sketch how these points of inquiry can lead to a more demonstrative analysis.
\end{itshape}
\newline\newline
Thank you for the comments and the additional suggestions to improve the manuscript. 
The main objective of the submitted manuscript is to construct a systematic approach to design low-thrust orbital maneuvers. 
It is motivated by the specific challenges that electric propulsion systems generate a minuscule level of thrust, and low-thrust orbital maneuvers rely on those effects accumulated over a long time period. 
It is quite cumbersome and challenging to apply numerical optimization tools or boundary value problem solvers successfully to low-thrust orbital design problems. 
We aim to provide more organized tools to mitigate such difficulties by explicitly considering the characteristics of the natural dynamics and utilizing geometric integrators exhibiting long-term numerical stability. 
The former is achieved by incorporating the control-free maneuvers obtained by the Poincare section in the invariant manifolds, and the latter is achieved by variational integrators. 
However, the proposed scheme does not guarantee any optimality in the constructed maneuvers. 
It is heuristically implied that the constructed orbital maneuver would require no-excessive control input, as it is essentially constructed by adjusting the control-free trajectory. 
In short,
\begin{itemize}
	\item[Pros] The proposed scheme utilizes the characteristics inherent in the three-body problem in constructing low-thrust orbital maneuvers in a systematic way. 
		This reduces the cumbersome process of guessing a \textit{good} set of initial parameters to make the direct application of optimization successful. 
	\item[Pros] The use of geometric integrators is particularly useful in capturing the long-term effects of low thrust in a computationaly efficient manner. 
	\item[Cons] While the proposed scheme generates orbital trajectories to the target with prescribed constraints, it does not guarantee any optimality. 
\end{itemize}
The conclusions section has been revised to summarize the above pros and cons of the proposed approach. 
\begin{itshape}
	Put the revised conclusions here.
\end{itshape}

### 2018-05-21 manuscript.tex
Merge changes from Dr. Lee to manuscript

% utilizing the combo of small satellite and low thrust propulsion offers new mission scenarios 
Recent developments in miniature electric propulsion and small satellites now offer the potential for new research oppportunities~\cite{folta2015,haque2013}.
The potential for more demanding missions places an even greater importance on the mission design to ensure that optimal trajectories satisfy mission requirements~\cite{folta2015,koon2011,ross2006,gomez2001}. 
There has been extensive research focused on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011,koon2011,ross2006}. 
Within the three-body problem, a spacecrafts feasible region of motion is constrained by its energy, or Jacobi integral.
The addition of low-thrust propulsion offers the potential of reduced transfer transit times and the ability to depart from the free motion trajectory to allow for increased transfer opportunities.
Frequently, insight into the problem or intuition on the part of the designer is required to determine initial conditions that will converge to the desired solution due to the inherent nonlinear and chaotic behavior of the three-body system~\cite{szebehely1967}.
Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
Variational integrators can allow for reduced computational effort without any loss of numerical stability or energy drift which exist in conventional integration schemes.

@@

A simulation comparing the variational integrator to a conventional Runge-Kutta method is presented in this section.
A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0.2 & 0\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
The variational integrator uses a range of step sizes between \SIrange{47.2}{4720}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
The step size of the variational integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference for this comparison.
Both integration schemes result in trajectories that are initially nearly identical.
\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the conventional integration scheme.
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
This work uses a second order integrator which has improved energy behavior in comparision to first order integration techniques without a significant increase in computational demand.
The variational integrator provides improved energy behavior over long integration times but does not guarantee state accuracy.
High order Runge-Kutta methods will likely provide a more accurate state trajectory, but at the expense of additional computational demand and total energy deviations.
However, for this preliminary analysis the variational integrator provides reduced computational demands, as compared with general-purpose numerical integrators with a similar level of numerical error, and is useful to characterize and explore the trade space in preliminary mission design scenarios.
In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
The presented results are based on the variational integrator with a fixed time step for numerical simulation and gradient computation.
In summary, this paper utilizes the variational integrator to capture the long-term effects of low-thrust devices accurately and efficiently.
The structural stability during numerical integration is critical for numerical optimization in the subsequent developments.

@@

\section{Conclusions}\label{sec:conclusion}
In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~sections is used to generate a transfer between planar periodic orbits in the three-body problem.
The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
The indirect optimal control formulation direclty solves for the state and costate vectors in the form of a two point boundary value problem which results in a lower number of design variables.
Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution and typically results in a much larger parameter set.
Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.
Using the reachability set on a \Poincare section reduces the dimensionality of the system and simplifies the analysis by avoiding the cost of a completely unmotivated exploration of the space.
However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
Furthermore, our approach relies heavily on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.

>>>>

% utilizing the combo of small satellite and low thrust propulsion offers new mission scenarios 
Recent developments in miniature electric propulsion and small satellites now offer the potential for new research opportunities~\cite{folta2015,haque2013}.
The potential for more demanding missions places an even greater importance on the mission design to ensure that optimal trajectories satisfy mission requirements~\cite{folta2015,koon2011,ross2006,gomez2001}. 
There has been extensive research focused on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011,koon2011,ross2006}. 
Within the three-body problem, a spacecrafts feasible region of motion is constrained by its energy, or Jacobi integral.
The addition of low-thrust propulsion offers the potential of reduced transfer transit times and the ability to depart from the free motion trajectory to allow for increased transfer opportunities.
Frequently, insight into the problem or intuition on the part of the designer is required to determine initial conditions that will converge to the desired solution due to the inherent nonlinear and chaotic behavior of the three-body system~\cite{szebehely1967}.
Conventional general-purpose Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, may fail to preserve geometric properties of the dynamic system numerically. 
Variational integrators can allow for reduced computational effort without any loss of numerical stability or energy drift which exist in conventional integration schemes.

@@

A simulation comparing the variational integrator to a conventional Runge-Kutta method is presented in this section.
A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0.2 & 0\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
The variational integrator uses a range of step sizes between \SIrange{47.2}{4720}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
The step size of the variational integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference for this comparison.
Both integration schemes result in trajectories that are initially nearly identical.
\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the conventional integration scheme.
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
This work uses a second order integrator which has improved energy behavior in comparison to first order integration techniques without a significant increase in computational demand.
The variational integrator is constructed according to a discrete version of Hamilton's principle, specifically for the given dynamic system, and it provides long-term structural stability in capturing the effects of low-thrust propulsion systems.
General purpose numerical integrators, such as adaptive stepsize Runge-Kutta methods may yield a more accurate state trajectory depending on the choice of integration parameters, but at the expense of additional computational demand and total energy deviations.
However, for this preliminary analysis the variational integrator provides reduced computational demands, as compared with general-purpose numerical integrators with a similar level of numerical error, and is useful to characterize and explore the trade space in preliminary mission design scenarios.
In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
The presented results are based on the variational integrator with a fixed time step for numerical simulation and gradient computation.
In summary, this paper utilizes the variational integrator to capture the long-term effects of low-thrust devices accurately and efficiently.
The structural stability during numerical integration is critical for numerical optimization in the subsequent developments.

@@

\section{Conclusions}\label{sec:conclusion}
In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~sections is used to generate a transfer between planar periodic orbits in the three-body problem.
The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
The indirect optimal control formulation directly solves for the state and costate vectors in the form of a two point boundary value problem which results in a lower number of design variables.
Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution and typically results in a much larger parameter set.
Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.
Using the reachability set on a \Poincare section reduces the dimensionality of the system and simplifies the analysis by avoiding the cost of a completely unmotivated exploration of the space.
However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.

Furthermore, our approach relies heavily on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.
